---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity <- read.csv("~/Reproductible research/activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
#Sum daily steps
daily_steps <- tapply(activity$steps,activity$date,sum, na.rm=TRUE)
#Draw histogram
hist(daily_steps, breaks = 20, main = "Total steps per day", xlab = "Number of steps", ylab = "Frequency")
#Report mean and median
mean1 <- mean(daily_steps)
median1 <- median(daily_steps)
```
### Mean and median steps per day
```{r}
print(mean1)
print(median1)
```

## What is the average daily activity pattern?

```{r}
#Average steps by interval
  interval_steps <- as.data.frame(cbind(tapply(activity$steps,activity$interval,mean, na.rm=TRUE),unique(activity$interval)))
  #Name columns
  colnames(interval_steps) <- c("Steps","Interval")  
  
#Draw histogram
  plot(interval_steps$Interval, interval_steps$Steps, type = "l", main = "Average steps througout day", xlab = "Time", ylab = "Steps")
#Find interval with most steps on average
  interval_steps$Interval[match(max(interval_steps$Steps),interval_steps$Steps)]
```

## Imputing missing values

### Missing variables are replaced by the average for that interval, based on the calculations of the previous question
```{r}
activity$steps <- ifelse(is.na(activity$steps),interval_steps$Steps[match(activity$interval,interval_steps$Interval)],activity$steps)
```

##Histogram of daily steps
```{r}
#Sum daily steps
daily_steps <- tapply(activity$steps,activity$date,sum, na.rm=TRUE)
#Draw histogram
hist(daily_steps, breaks = 20, main = "Total steps per day", xlab = "Number of steps", ylab = "Frequency")
#Report mean and median
mean2 <- mean(daily_steps)
median2 <- median(daily_steps)
```
## Mean and median steps with filled data
```{r}
  mean2
  median2
```
## Difference with filled data
```{r}
mean2-mean1
median2-median1
```

## Are there differences in activity patterns between weekdays and weekends?

### Create weekday variable
```{r}
activity$weekday <- ifelse(weekdays(as.Date(activity$date))=="Saturday"|weekdays(as.Date(activity$date))=="Sunday","Weekend","Weekday")
```

```{r}
#Average steps by interval
  weekday_interval_steps <- as.data.frame(cbind(tapply(activity$steps[activity$weekday=="Weekday"],activity$interval[activity$weekday=="Weekday"],mean, na.rm=TRUE),unique(activity$interval[activity$weekday=="Weekday"])))
  #Name columns
  colnames(weekday_interval_steps) <- c("Steps","Interval")  
#Average steps by interval
  weekend_interval_steps <- as.data.frame(cbind(tapply(activity$steps[activity$weekday=="Weekend"],activity$interval[activity$weekday=="Weekend"],mean, na.rm=TRUE),unique(activity$interval[activity$weekday=="Weekend"])))
  #Name columns
  colnames(weekend_interval_steps) <- c("Steps","Interval")
  
```    
#Draw histogram
```{r}
#Weekday plot
  plot(weekday_interval_steps$Interval, weekday_interval_steps$Steps, type = "l", main = "Average steps througout weekday", xlab = "Time", ylab = "Steps")
#Weekend plot
  plot(weekend_interval_steps$Interval, weekend_interval_steps$Steps, type = "l", main = "Average steps througout weekend", xlab = "Time", ylab = "Steps")
``` 